---
title: "Estimate Correlation"
author: "Kristen Hunter"
date: "2/12/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(PUMP)
source(here::here('code', 'estimate_power_with_simulation.R'))
knitr::opts_chunk$set(echo = TRUE)
```

# Correlation estimation strategy

- For $s = 1, \dots, S$:
  - Input DGP parameters, including assumed correlation $\rho$.
  - Generate full simulated data. 
  - Generate treatment assignments and observed data.
  - Calculate $t$ statistics from regressions for each outcome. So for $M$ outcomes, we have a $M$-vector of test statistics.
- Returns a matrix of test statistics $S \times M$.

Next:

- Calculate correlation of matrix of test statistics.
- Take upper triangle of matrix and take average.
- Returns a single mean estimated correlation.

The following function returns the $S \times M$ matrix of test statistics:
```{r}
get_rawt <- function(d_m, model.params.list, Tbar, n.sims = 100)
{
  M <- model.params.list$M
  rawt.all <- matrix(NA, nrow = n.sims, ncol = M)
  dgp.params.list <- PUMP::convert_params(model.params.list)
  
  # number of simulations
  for(s in 1:n.sims)
  {
    if (s %% 20 == 0){ message(paste0("Now processing simulation ", s, " of ", n.sims)) }
    
    # generate simulated data
    samp.full <- PUMP::gen_full_data(dgp.params.list)
    S.id <- samp.full$ID$S.id
    D.id  <- samp.full$ID$D.id
    
    # generate treatment assignments
    T.x <- PUMP::gen_T.x(d_m = d_m,
                         S.id = S.id, D.id = D.id,
                         nbar = dgp.params.list$nbar,
                         Tbar = 0.5)
    
    # convert full data to observed data
    samp.obs <- samp.full
    samp.obs$Yobs <- PUMP::gen_Yobs(samp.full, T.x)
    
    # calculate t statistics
    dat.all <- makelist_samp(samp.obs, T.x)
    rawpt.out <- get_rawpt(dat.all, d_m = d_m, model.params.list = model.params.list)
    rawt <- sapply(rawpt.out[['rawpt']], function(s){ return(s[['tstat']])})
    rawt.all[s,] <- rawt
  }
  
  return(rawt.all)
}
```

The following function takes in the matrix of test statistics and returns the estimated correlation.

```{r}
get_cor <- function(rawt.all)
{
  
  # calculate correlation
  cor.tstat <- cor(rawt.all)
  est.cor <- cor.tstat[lower.tri(cor.tstat)]
  mean.est.cor <- mean(est.cor)

  return(mean.est.cor)
}
```


# Simulation results

The following sets up the design and DGP parameters.

```{r, echo = FALSE}
model.params.list <- list(
  M = 3                             # number of outcomes
  , J = 30                          # number of schools
  , K = 10                          # number of districts
  , nbar = 50                       # number of individuals per school
  , rho.default = 0.5               # default rho value (optional)
  ################################################## impact
  , MDES = 0.125                    # minimum detectable effect size      
  ################################################## level 3: districts
  , numCovar.3 = 1                  # number of district covariates
  , R2.3 = 0.1                      # percent of district variation
  # explained by district covariates
  , ICC.3 = 0.2                     # district intraclass correlation
  , omega.3 = 0.1                   # ratio of district effect size variability
  # to random effects variability
  ################################################## level 2: schools
  , numCovar.2 = 1                  # number of school covariates
  , R2.2 = 0.1                      # percent of school variation
  # explained by school covariates
  , ICC.2 = 0.2                     # school intraclass correlation	
  , omega.2 = 0.1                   # ratio of school effect size variability
  # to random effects variability
  ################################################## level 1: individuals
  , numCovar.1 = 1                  # number of individual covariates
  , R2.1 = 0.1                      # percent of indiv variation explained
  # by indiv covariates
)
Tbar <- 0.5
```


```{r}
# simulation parameters
n.sims <- 10
d_m.list <- c('d2.1_m2fr', 'd2.2_m2rc', 'd3.1_m3rr2rr', 'd3.2_m3rr2rc')
rho.list <- c(0.1, 0.5, 0.8)
```

```{r echo = FALSE}
# store results
out.data <- NULL

for(d_m in d_m.list)
{
  print(paste('Design:', d_m))
  for(rho in rho.list)
  {
    print(paste('Rho:', rho))
    rawt.all <- get_rawt(
      d_m = d_m, model.params.list = model.params.list, Tbar = Tbar, n.sims = n.sims
    )
    est.cor <- get_cor(rawt.all)
  
    cor.data <- data.frame(
      d_m = d_m,
      input.rho = rho,
      output.rho = est.cor,
      n.sims = n.sims
    )
    out.data <- rbind(out.data, cor.data)
  }
}

print(out.data)
```


  
  

