---
title: "Monte Carlo Simulation Code"
author: "Kristin Porter and Deni Chen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

This code generates the Monte Carlo simulations for validating methods in the paper (table C.3) and calls items from I:\Multiplicity\Archive\Domino Copy\ECmethods\R.


Clear everything from memory
```{r clear_memory}
rm(list=ls())
```

Set up: install and load libraries and source functions
```{r source}
#source("libraries.install.R")

library(RcppEigen)
library(snow)
library(lme4)

source("http://bioconductor.org/biocLite.R")
biocLite("multtest")
library(multtest)

source("gen.blocked.data.R")
source("adjust.WY.R")
source("functions.R")

#funct object needs to be assigned before sourcing "make.model.R"
funct<-"fixfastLm"; mod.type<-"fixed"
source("make.model.R")
```


Specify parameters
* rho         = Spearman's correlation determines the strength and direction of the monotonic relationship between two variables rather than the                 strength and direction of the linear relationship between your two variables (Pearson's correlation)    
* ncl         = number of clusters set at 24 because that is the max cpus on datalab
* procs       = a vector of strings for adjustment procedures
* M           = number of tests/domains/outcomes
* MDES        = minimum detectable effect size, vector length M 
* p.j.range   = vector of minimum and maximum probabilities of being assigned to treatment, across all sites
* S           = number of samples for power calc
* B           = number of permutations for WY
* J           = number of blocks
* n.j         = number of observations per block 
* theta       = MxM matrix of correlations between residuals in Level 2 model outcomes under no treatment and Level 2 effects 
* omega       = effect size variability, between 0 and 1, 0 if no variation in effects across blocks, vector length M
* Gamma.00    = grand mean outcome w/o treat, held 0, vector length M 
* sig.sq      = vector length M, held at 1 for now
* alpha       = the significance level, 0.05 usually 
* ICC         = a number, intraclass correlation; 0 if fixed model  
* R2.1        = R squared for mth level 1 outcome by mth level 1 covar          
* R2.2        = R squared for mth level 2 outcome by mth level 1 covar          
* rho.0_lev1  = MxM matrix of correlations for Level 1 residuals in models of outcomes under no treatment                   
* rho.0_lev2  = MxM matrix of correlations for Level 2 residuals in models of outcomes under no treatment                   
* rho.1_lev2  = MxM matrix of correlations for Level 2 effects        
* DDMDES      = desired minimum detectable effect size, number 



```{r specifications}
rho<-0.8

ncl<-8

procs<-c("Bonferroni", "BH", "Holm", "WY") 
M<-6
MDES<-rep(0.125, M)
p.j.range<-c(0.5,0.5)
#S=2000
S=3
B=300   #10000
J=20;n.j=100
theta<-matrix(0,M,M); diag(theta)<-0; omega <- rep(0,M)
Gamma.00<-rep(0,M);sig.sq<-rep(1,M); alpha<-0.05
ICC<-rep(0, M); R2.2<-rep(0,M); R2.1<-rep(0, M)

rho.0_lev1<-matrix(rho,M,M); diag(rho.0_lev1)<-1 
rho.0_lev2<-matrix(0.5,M,M); diag(rho.0_lev2)<-1
rho.1_lev2<-matrix(0.5,M,M); diag(rho.1_lev2)<-1


#pastes together specifications into a file name
simname<-paste0("M", M, "n.j", n.j, "J", J, "ICC", ICC[1], "MDES", MDES[1], "rho", rho, "_S", S, "B", B, "_R2.1", R2.1[1],"_R2.2",R2.2[1], mod.type, "_sim.Rda")
```

Simulate data
```{r simulate_data}
startingtime<-Sys.time()

simpwr<-est.power(procs=procs, M=M, DMDES=MDES, n.j=n.j, J=J, rho.0_lev1=rho.0_lev1, 
                  rho.0_lev2=rho.0_lev2, rho.1_lev2=rho.1_lev2, theta=theta, ICC=ICC, 
                  alpha=alpha, Gamma.00=Gamma.00, sig.sq=sig.sq, p.j.range=p.j.range, 
                  R2.1=R2.1, R2.2=R2.2, check=FALSE, omega=omega, funct=mod.type, S=S, ncl=ncl, B=B,maxT=FALSE)     

endingtime<-Sys.time()

print(paste("Run with S=", S, " and B =", B, " with ncl=", ncl, " took ", difftime(endingtime,startingtime, units = "secs"), " seconds. Started at", startingtime, "Ended at ", endingtime))

#what is maxT and check?

#save(simpwr, file=simname)
```


Make replica of this RMD into an R Script (.R file to be sourced into another pgm)
```{r create_RMD}
#library(knitr)
#purl("MonteCarloSimulation.Rmd", output = "MonteCarloSimulation.R", documentation = 2)
```
