---
title: "Profiling code speed"
date: 02/05/2019
output: pdf_document
---

## Loading the blockrct2_power.R file

```{r}
source("blockrct2_power.R")
#library(pum)
```
## Parameters for testing power function en masse

The initial conditions for power are set as below. These are initial conditions set of a vector of length 5 that would allow us to use expand grid to test out all different combinations of scenarios.

```{r}
M <- c(5:10)
MDES <- seq(from = 0.125, to = 0.1, length.out = 5 )
Aimpacts <- seq(from = 5, to = 1, length.out = 5)
J <- seq(10, 50, length.out = 5)
n.j <- seq(50,100, length.out = 5)
R2.1 <- seq(0.4, 0.9, length.out = 5)
p <- 0.5
alpha <- 0.05
numCovar.1 <- c(5:10)
```

## Test Cases before implmenting optimization

### Summary 

The key drivers of computational time are Westfall-Young Single Step and Westfall-Young Step Down procedures within the power function.

* Increasing the baseline parameters other than sample number (snum) are not going to significantly change the computational time.

* Increasing snum, however, affects computational time significantly to as much as 3 minutes per power calculation should we want to calculate base on set initial conditions.

* Parallelization appears to be working as WestFall-Young Step Down and Westfall-Young Single Step takes about the same time. By code design, WYSD has an additional loop structure going compared to WYSS.

### Test case 1 for power

An initial test case with a **baseline** parameters set as below. The total elapsed time for the computation is **5.4** seconds. 
The majority of the time is spent on apply function using multtest's built in function to find adjustments for all the MTPs except Westfall-Young at 2.2 seconds. The rest is split in over 1 seconds each with Westfall-Young SD and Westfall-Young Single Step. Sample number for Westfall-Young is set at 10.

* M = 5
* MDES = 0.125
* Ai = 3
* J = 10
* n.j = 10
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 10
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 5, MDES = 0.125, Ai = 3 , J = 10 , n.j = 10, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 10, updateProgress = NULL)
```

### Test case 2 for Power

Here the **baseline** parameters of the harmonic mean is made 10 times compared to in test case 1. The number of M outcomes is made twice the original set amount of 5, too. The total elapsed time for the computation is ~ **8** seconds. 
Again, the majority of the time is spent on apply function using multtest's built in function to find adjustments for all the MTPs except Westfall-Young at ~ 3 seconds. We see increases in both Westfall-Young SD and Westfall-Young SS, too with more time allocated towards WYSD.

* M = 10
* MDES = 0.125
* Ai = 3
* J = 10
* n.j = 100
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 10
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 10, MDES = 0.125, Ai = 3 , J = 10 , n.j = 100, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 10, updateProgress = NULL)
```

### Test case 3 for Power

In test case 3, we go back to the baseline conditions for all the parameters except the number of sample as it applies to Westfall-Young. The target number of samples stands at 1,000 but, currently, we have increased it by **10** times to **1,00** from **10**.

This leads to a significant increase in the computation time up to **22** seconds from the original **5.4** seconds.We are seeing an increase of about **4** times. The allocation across programming pieces have also shifted towards Westfall-Young procedures taking the balk of it at **10** and **9** seconds each with not much in the multtest procedure. 

* M = 5
* MDES = 0.125
* Ai = 3
* J = 10
* n.j = 10
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 100
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 10, MDES = 0.125, Ai = 3 , J = 10 , n.j = 100, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 100, updateProgress = NULL)
```

### Test case 4 for Power

In test case 4, we have increased some of the baseline parameters such as M, J and n.j. Additionally, we doubled the number of samples for Westfall-Young from 100 to 200, a doubling. With this, we are seeing a further increase in computation time of up to **37** seconds with a non-linear increase in Westfall-Young procedure of up to **16** seconds each.

* M = 10
* MDES = 0.125
* Ai = 3
* J = 50
* n.j = 100
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 200
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 10, MDES = 0.125, Ai = 3 , J = 10 , n.j = 100, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 200, updateProgress = NULL)
```


### Test case 5 for Power

In test case 5, we have maintained the same parameters as before except the sample for Westfall-Young is increased from 200 to 500. With this, we see a significant increase in computation time of up to **86** (i.e 1 minute and 26) seconds. The majority of the time elapsed increase could be contributed as before to Westfall-Young procedures each clocking **40** seconds.

* M = 10
* MDES = 0.125
* Ai = 3
* J = 50
* n.j = 100
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 500
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 10, MDES = 0.125, Ai = 3 , J = 50 , n.j = 100, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 500, updateProgress = NULL)
```

### Test case 6 for Power

In test case 6, we change the number of clusters used to run the power calculation from 2 to 4. The goal was to see whether with increased cluster, we see a decline in computational time. On the contrary, we don't. The computational time increases in fact by **2** seconds to **88** (i.e 1 minute 20 seconds) seconds. More runs are needed to confirm whether there is a statistical fluke or not.

* M = 10
* MDES = 0.125
* Ai = 3
* J = 50
* n.j = 100
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 500
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 10, MDES = 0.125, Ai = 3 , J = 50 , n.j = 100, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 500, updateProgress = NULL)
```


### Test case 7 for Power

Number of clusters is back to 2 but, the sample number for Westfall-Young is doubled further to the minimum a thousand that is required for a good run of Westfall-Young. All the other parameters remain the same. Doubling than the number of sample size leads to more than a doubling of computation time at **3** minutes. 

* M = 10
* MDES = 0.125
* Ai = 3
* J = 50
* n.j = 100
* R2.1 = 0.5
* p = 0.5
* alpha = 0.05
* numCovar.1 = 5
* numCovar.2 = NULL
* ICC = NULL
* tnum = 10000
* snum = 1000
* updateProgress = NULL

```{r}
power.blockedRCT.2(M = 10, MDES = 0.125, Ai = 3 , J = 50 , n.j = 100, R2.1 = 0.5 , p = 0.5 , alpha = 0.05 , numCovar.1 = 5 , numCovar.2 = NULL, ICC = NULL, tnum = 10000, snum = 1000, updateProgress = NULL)
```










