---
title: "Estimate Correlation"
author: "Kristen Hunter"
date: "2/12/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(PUMP)
source(here::here('code', 'estimate_power_with_simulation.R'))
knitr::opts_chunk$set(echo = TRUE)

run.name = '20220218_23'
```

# Correlation estimation strategy

- For $s = 1, \dots, S$:
  - Input DGP parameters, including assumed correlation $\rho$.
  - Generate full simulated data. 
  - Generate treatment assignments and observed data.
  - Calculate $t$ statistics from regressions for each outcome. So for $M$ outcomes, we have a $M$-vector of test statistics.
- Returns a matrix of test statistics $S \times M$.

Next:

- Calculate correlation of matrix of test statistics.
- Take upper triangle of matrix and take average.
- Returns a single mean estimated correlation.

# Code

The following function returns the $S \times M$ matrix of test statistics:
```{r}
get_rawt <- function(d_m, model.params.list, Tbar, n.sims = 100)
{
  M <- model.params.list$M
  rawt.all <- matrix(NA, nrow = n.sims, ncol = M)
  dgp.params.list <- PUMP::convert_params(model.params.list)
  
  # number of simulations
  for(s in 1:n.sims)
  {
    if (s %% 20 == 0){ message(paste0("Now processing simulation ", s, " of ", n.sims)) }
    
    # generate simulated data
    samp.full <- PUMP::gen_full_data(dgp.params.list)
    S.id <- samp.full$ID$S.id
    D.id  <- samp.full$ID$D.id
    
    # generate treatment assignments
    T.x <- PUMP::gen_T.x(d_m = d_m,
                         S.id = S.id, D.id = D.id,
                         nbar = dgp.params.list$nbar,
                         Tbar = 0.5)
    
    # convert full data to observed data
    samp.obs <- samp.full
    samp.obs$Yobs <- PUMP::gen_Yobs(samp.full, T.x)
    
    # calculate t statistics
    dat.all <- makelist_samp(samp.obs, T.x)
    rawpt.out <- get_rawpt(dat.all, d_m = d_m, model.params.list = model.params.list)
    rawt <- sapply(rawpt.out[['rawpt']], function(s){ return(s[['tstat']])})
    rawt.all[s,] <- rawt
  }
  
  return(rawt.all)
}
```

The following function takes in the matrix of test statistics and returns the estimated correlation.

```{r}
get_cor <- function(rawt.all)
{
  
  # calculate correlation
  cor.tstat <- cor(rawt.all)
  est.cor <- cor.tstat[lower.tri(cor.tstat)]
  mean.est.cor <- mean(est.cor)

  return(mean.est.cor)
}
```




# Simulation Results

```{r, echo = FALSE}
results = readRDS(here::here(paste0('cor_results_', run.name , '.rds')))
kable(results, digits = 3)
```