---
title: "Summary of validation results"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---


```{r knit.options, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, warning = FALSE, message = FALSE, echo = FALSE,
  fig.height = 7
)
opts <- options(knitr.kable.NA = "--")
```

```{r setup}
library(here)
library(kableExtra)
library(magrittr) 
library(PUMP)
library(pander)
library(tibble)
library(tidyr)
# default tables to round to 3 digits
panderOptions('round', 3)
source(here::here("code", "misc.R"))
```

# Introduction and notation

\textbf{Simulated power intervals}.
The Monte Carlo simulations produce a 95\% confidence interval, and we check whether that interval contains the `PUMP` estimates.
In other words, we check if $\hat{p}_{pump}$ is within $\hat{p}_{sim} \pm 1.96 \sqrt{0.25/5000}$.

\textbf{Simulated power point estimates}.
We compare the point estimates from the simulations to `PUMP`.
To do so, we calculate the absolute difference between the power estimates from the simulations and from `PUMP`:
$$ b_{sim} = \lvert \hat{p}_{sim} - \hat{p}_{pump} \rvert.$$

\textbf{PowerUpR power point estimates}.
We compare the point estimates from the `PowerUpR` to `PUMP`.
This comparison is only conducted for individual, unadjusted power.
Similar to the last metric, we calculate the absolute difference between the power estimates from `PowerUpR` and from `PUMP`:
$$ b_{pow} = \lvert \hat{p}_{pow} - \hat{p}_{pump} \rvert.$$



\textbf{PUMP mdes point estimates}.
We summarize the MDES performance by calculating the absolute difference between the `PUMP` estimate and the target MDES.
$$ b_{mdes} = \lvert \hat{m}_{pump} - \hat{m}_{target} \rvert.$$

```{r}
# extract and summarize results for a particular d_m
get.d_m.power.results <- function(d_m, power.files)
{
  results.files <- grep(d_m, power.files, value = TRUE)
  
  # stack results
  d_m.results <- NULL
  for(file in results.files)
  {
    print(file)
    results <- readRDS(here::here('output', file))

    results.wide <- results %>%
      pivot_wider(names_from = c(method, value.type), values_from = value)
    
    d_m.results <- rbind(d_m.results, results.wide)
  }

  row.names(d_m.results) <- NULL
  # d_m.results %<>% mutate(
  #   cover =
  #     Sim_ci_lower <= PUMP_adjusted_power &
  #     Sim_ci_upper >= PUMP_adjusted_power,
  #   bias.sim = abs(Sim_adjusted_power - PUMP_adjusted_power) / PUMP_adjusted_power,
  #   bias.pow = abs(PowerUp_adjusted_power - PUMP_adjusted_power) / PUMP_adjusted_power
  # )
  d_m.results %<>% mutate(
    cover =
      Sim_ci_lower <= PUMP_adjusted_power &
      Sim_ci_upper >= PUMP_adjusted_power,
    bias.sim = abs(Sim_adjusted_power - PUMP_adjusted_power),
    bias.pow = abs(PowerUp_adjusted_power - PUMP_adjusted_power)
  )
  d_m.results$d_m <- d_m
  
  # append information about whether we expect results to agree
  d_m.results$agree <- TRUE
  if(d_m %in% c('d2.1_m2fc', 'd2.1_m2ff'))
  {
    d_m.results$agree[d_m.results$ICC.2 > 0] <- FALSE
  }

  return(d_m.results)
}
```


```{r}
# find files
all.files <- list.files(here::here('output'))
# restrict to relevant files
power.files <- grep('power', all.files, value = TRUE)
# subset into WY and not WY
main.power.files <- grep('__B', power.files, value = TRUE)
wy.power.files <- power.files[!(power.files %in% main.power.files)]

d_m.list <- c(
  "d2.1_m2fc", "d2.1_m2ff", "d2.1_m2fr",
  "d2.2_m2rc", "d3.1_m3rr2rr", "d3.2_m3ff2rc", "d3.2_m3rr2rc", "d3.3_m3rc2rc"
)

all.power.results <- NULL
for(d_m in d_m.list)
{
  print(d_m)
  d_m.results <- get.d_m.power.results(d_m, main.power.files)
  all.power.results <- rbind(all.power.results, d_m.results)
}

```

```{r}
# summarize coverage
cover.results <- ddply(
  all.power.results[all.power.results$agree,], c('d_m', 'MTP', 'power_type'),
  summarise,
  cover = mean(cover)
)
```

```{r}
# summarize bias
bias.results <- ddply(
  all.power.results[!is.na(all.power.results$cover) & all.power.results$cover,],
  c('d_m', 'MTP', 'power_type'),
  summarise,
  mean.b.sim = mean(bias.sim),
  mean.b.pow = mean(bias.pow[agree == TRUE], na.rm = TRUE)
)
# clean up for table printing
bias.results$mean.b.pow[is.nan(bias.results$mean.b.pow)] = NA
```

# Summary of validation coverage results

```{r, results = 'asis'}
for(scenario in d_m.list)
{
  table.data <- dplyr::filter(cover.results, d_m == scenario)
  table.data <- table.data  %>%
    dplyr::arrange(MTP, power_type)
  
  print(kable(table.data, booktabs = TRUE, row.names = FALSE, digits = 3,
              linesep = c("", "", "", "", "", "", "\\addlinespace")))
}
```

\clearpage
\newpage

# Coverage discrepancies

We summarize below the scenarios where the simulation intervals do not cover the PUMP value.
For brevity, we only display results for Bonferroni adjustments, because 
```{r, results = 'asis'}
# extract out problem rows
problems <- all.power.results[
  !is.na(all.power.results$cover) &
  all.power.results$cover != 1 &
  all.power.results$agree,
]
problems <- problems %>%
  dplyr::select(d_m, MTP, power_type, omega.2, omega.3, ICC.2, ICC.3, 
         PUMP_adjusted_power, PowerUp_adjusted_power, Sim_adjusted_power,
         Sim_ci_lower, Sim_ci_upper) %>%
  dplyr::rename(
         type = power_type,
         pump = PUMP_adjusted_power,
         pow = PowerUp_adjusted_power,
         sim = Sim_adjusted_power,
         low = Sim_ci_lower,
         up = Sim_ci_upper)

problem.table <- data.frame(problems)

for(scenario in d_m.list)
{
  problem.table.data <- dplyr::filter(problem.table, d_m == scenario, MTP == 'BF')
  
  if(nrow(problem.table.data) > 0)
  {
    print(kable(problem.table.data, booktabs = TRUE, row.names = FALSE, digits = 3))
  }
}
```

\clearpage
\newpage


# Summary of validation "bias" results

```{r, results = 'asis'}
for(scenario in d_m.list)
{
  table.data <- dplyr::filter(bias.results, d_m == scenario)
  table.data <- table.data  %>%
    dplyr::arrange(MTP, power_type)
  
  print(kable(table.data, booktabs = TRUE, row.names = FALSE, digits = 3,
              linesep = c("", "", "", "", "", "", "\\addlinespace")))
}
```

\newpage
# Collapsed Summaries

```{r}
bias.results.summary <- ddply(
  bias.results, 'd_m',
  summarise,
  avg.b.sim = mean(mean.b.sim),
  max.b.sim = max(mean.b.sim),
  mean.b.pow = mean(mean.b.pow, na.rm = TRUE)
)
colnames(bias.results.summary) <- c('d_m', 'mean.b.sim', 'max.b.sim', 'mean.b.pow')
kable(bias.results.summary, booktabs = TRUE, row.names = FALSE, digits = 3)
# kable(bias.results.summary, booktabs = TRUE, row.names = FALSE, digits = 3,
# format = 'latex')
```

# WY Summary

```{r}
wy.power.results <- NULL
for(d_m in d_m.list)
{
  d_m.results <- get.d_m.power.results(d_m, wy.power.files)
  wy.power.results <- rbind(wy.power.results, d_m.results)
}
wy.power.results <- wy.power.results[wy.power.results$MTP %in% c('WY-SS', 'WY-SD'),]
```


```{r}
wy.cover.results <- ddply(
  wy.power.results, c('d_m', 'MTP', 'power_type'),
  summarise,
  cover = mean(cover)
)
```


```{r}
# summarize bias
wy.bias.results <- ddply(
  wy.power.results[!is.na(wy.power.results$cover) & wy.power.results$cover,],
  c('d_m', 'MTP', 'power_type'),
  summarise,
  mean.b.sim = mean(bias.sim)
)
```

```{r}
wy.bias.results.summary <- ddply(
  wy.bias.results, c('d_m', 'MTP'),
  summarise,
  avg.b.sim = mean(mean.b.sim),
  max.b.sim = max(mean.b.sim)
)
colnames(wy.bias.results.summary) <- c('d_m', 'MTP', 'mean.b.sim', 'max.b.sim')
kable(wy.bias.results.summary, booktabs = TRUE, row.names = FALSE, digits = 3)
# kable(wy.bias.results.summary, booktabs = TRUE, row.names = FALSE, digits = 3,
# format = 'latex')
```

# MDES summary


```{r}
# find files
all.files <- list.files(here::here('output'))
# restrict to relevant files
mdes.files <- grep('mdes', all.files, value = TRUE)
mdes.files <- grep('5000_S', mdes.files, value = TRUE)
```

```{r}
# extract and summarize results for a particular d_m
get.d_m.mdes.results <- function(d_m,  mdes.files)
{
  results.files <- grep(d_m, mdes.files, value = TRUE)
  # remove WY files
  mdes.file <- grep('__B', results.files, value = TRUE)
  
  results <- readRDS(here::here('output', mdes.file))

  results$abs.rel.bias <- abs(results$`Adjusted MDES` - results$`Target MDES`) /  results$`Target MDES`

  return(results)
}
```


```{r}
all.mdes.results <- NULL

for(d_m in d_m.list)
{
  d_m.results <- get.d_m.mdes.results(d_m, mdes.files)
  all.mdes.results <- rbind(all.mdes.results, d_m.results)
}
```

```{r}
kable(all.mdes.results, booktabs = TRUE, row.names = FALSE, digits = 3)
```

```{r}
mdes.results.summary <- ddply(
  all.mdes.results, 'd_m',
  summarise,
  mean.b.mdes = mean(abs.rel.bias, na.rm = TRUE)
)
kable(mdes.results.summary, booktabs = TRUE, row.names = FALSE, digits = 3)
# kable(mdes.results.summary, format = 'latex', booktabs = TRUE, row.names = FALSE, digits = 3)
```



