---
title: "Validate_power_mdes_sample_blocked_i1_2c"
author: "Kristin Porter & Zarni Htet"
output:
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: monochrome
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float: yes
  github_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    toc: yes
---
# Purpose

This document lays out validation steps & results for estimating statistical power, minimum detectable effect size (MDES) and sample size for 2-Level Block RCT with constant effects experimental design. We compare results of the derived methods against Monte Carlo Simulations & the PowerUp R package on power, mdes and sample size. The methods are derived by Kristin Porter as outlined in the paper/s here (Add reference section). 

### Set Up

In this section, we set up the required R packages and source files to run the rest of the Validation document.

```{r libraries, warnings = FALSE, message = FALSE}
library(RcppEigen) # Linear algebra calculations
library(snow)      # Parallelization and use of multiple clusters 
library(lme4)      # Linear mixed effect models
library(PowerUpR)  # Comparison with another package that has some limitations
library(multtest)  # Multiple Testing Procedures package
library(tictoc)    # For timing of the scripts
library(here)      # For relative path sourcing
library(pum)       # For Power Under Multiplicity Package
library(dplyr)     # For combining different dataframes
library(pander)    # For good table format
```

```{r scriptFiles, warnings = FALSE, message = FALSE}
source(here::here("Validation/Simulations", "adjust.WY.R")) # MTP Westfall-Young
source(here::here("Validation/Simulations", "gen_data_dist_2_levels.R")) # data generating functions for 2 level block and cluster
source(here::here("Validation/Simulations", "power_mcs_2_levels.R")) # Estimate power for each of the designs by running data generating function S sample times and fitting models to estimate power.
source(here::here("Validation/Simulations", "validate_power_mdes_sample_blocked_i1_2c.R")) # Wrapping and creating 
source(here::here("Validation/Simulations", "misc.R")) # For coloring texts and other purposes
```

# Power Validation

In this section, we validate power results for different definitions of power, __minimum__, __individual__ and __complete__ for three multiple adjustment multiple methods, __Bonferroni__, __Benjamini Hocheberg__ and __Holm__.

Each subsection of the power validation below is a variation of correlation between outcomes, __rho__ & the ratio of __false nulls__ in M outcomes. For more details on input parameters, please refer to the paper or the Appendix section at the bottom of this markdown.

### n.j = 100, M = 3,numFalse = M, J = 20, rho = (0.5, 0.5, 0.5), MDES = (0.125, 0.125, 0.125), R2 = (0,0,0)

If you are previewing this in html, please click on the code section to review more detailed parameters.

```{r rho0.5numFalseMSpecs01, warning = FALSE, message = FALSE}
# ICC needs to be zero for constant and fixed effets
design <- c("Blocked_i1_2c")
procs <- c("Bonferroni", "BH", "Holm") 
M <- 3 
MDES <- rep(0.125, M) 
rho <- c(0.5)
ncl <-2 
S = 2000 # Number of samples for Monte Carlo Simulation.
B = 2 # Number of samples for WestFall-Young. The equivalent is snum in our new method.
tnum = 10000 # Number of test statistics (samples) for all procedures other than Westfall-Young. The default is set at 10,000.
J = 20 
n.j = 100 
p.j <- 0.5 
p.j.range <- rep(0.5,J) 
theta <- matrix(0,M,M) 
diag(theta) <- 0 
omega <- rep(0,M) 
Gamma.00 <- rep(0,M) 
sig.sq <- rep(1,M)
alpha <- 0.05 
ICC <- rep(0, M) 
R2.2 <- rep(0,M)  
R2.1 <- rep(0,M) 
maxT <- FALSE 
check <- FALSE # Usually for checking the code. Do not need to  
storage_path <- "./Validation/data"
validation_name <- "Blocked_i1_2c"
runSim <- TRUE
runPump <- TRUE
```

`r colorize("Individual Statistical power is estimated to be around 0.8. Please check the table below for estimations of other definitions of statistical power. Across all power definitions, estimation results across the __PUMP__ package, __Monte Carlo Simulation__ results and __PowerUp__ package are about the same as detailed below.", "green")`

`r colorize("We will be adding Westfall-Young step down adjustment later on.", "red")`

```{r rho0.5numFalseMRun01, warning = FALSE, message = FALSE}
blocked_i1_2c_compare_results <- validate_power_blocked_i1_2cfr(rho = rho, ncl = ncl, procs = procs, design = design, 
                                                                M = M, MDES = MDES,
                                                                p.j = p.j,p.j.range = p.j.range, S = S, B = B, 
                                                                tnum = tnum, J = J, 
                                                                n.j = n.j, theta = theta, omega = omega, 
                                                                Gamma.00 = Gamma.00, sig.sq = sig.sq,
                                                                alpha = alpha, ICC = ICC, R2.2 = R2.2, R2.1 = R2.1, 
                                                                maxT = maxT, rho.0_lev1 = rho.0_lev1,
                                                                rho.0_lev2 = rho.0_lev2, rho.1_lev2 = rho.1_lev2, 
                                                                check = check, storage_path = storage_path, 
                                                                validation_name = validation_name, runSim = runSim, 
                                                                runPump = runPump)

pander::pandoc.table(blocked_i1_2c_compare_results, paste0("PowerSimulation_", "rho_", rho,"_falseNull_",M), 
                     style = "grid", split.tables = 100)
print(blocked_i1_2c_compare_results)
```

### n.j = 75, M = 3, numFalse = M, J = 20, rho = (0.5, 0.5, 0.5), MDES = (0.125, 0.125, 0.125), R2 = (0,0,0)

```{r rho0.5numFalseMSpecs02, warning = FALSE, message = FALSE}
# ICC needs to be zero for constant and fixed effets
design <- c("Blocked_i1_2c")
procs <- c("Bonferroni", "BH", "Holm") 
M <- 3 
MDES <- rep(0.125, M) 
rho <- c(0.5)
ncl <-2 
S = 2000 # Number of samples for Monte Carlo Simulation.
B = 2 # Number of samples for WestFall-Young. The equivalent is snum in our new method.
tnum = 10000 # Number of test statistics (samples) for all procedures other than Westfall-Young. The default is set at 10,000.
J = 20 
n.j = 75 
p.j <- 0.5 
p.j.range <- rep(0.5,J) 
theta <- matrix(0,M,M) 
diag(theta) <- 0 
omega <- rep(0,M) 
Gamma.00 <- rep(0,M) 
sig.sq <- rep(1,M)
alpha <- 0.05 
ICC <- rep(0, M) 
R2.2 <- rep(0,M)  
R2.1 <- rep(0,M) 
maxT <- FALSE 
check <- FALSE # Usually for checking the code. Do not need to  
storage_path <- "./Validation/data"
validation_name <- "Blocked_i1_2c"
runSim <- TRUE
runPump <- TRUE
```

`r colorize("Individual Statistical power is estimated to be around 0.8. Please check the table below for estimations of other definitions of statistical power. Across all power definitions, estimation results across the __PUMP__ package, __Monte Carlo Simulation__ results and __PowerUp__ package are about the same as detailed below.", "green")`

`r colorize("We will be adding Westfall-Young step down adjustment later on.", "red")`

```{r rho0.5numFalseMRun02, warning = FALSE, message = FALSE}
blocked_i1_2c_compare_results <- validate_power_blocked_i1_2cfr(rho = rho, ncl = ncl, procs = procs, design = design, 
                                                                M = M, MDES = MDES,
                                                                p.j = p.j,p.j.range = p.j.range, S = S, B = B, 
                                                                tnum = tnum, J = J, 
                                                                n.j = n.j, theta = theta, omega = omega, 
                                                                Gamma.00 = Gamma.00, sig.sq = sig.sq,
                                                                alpha = alpha, ICC = ICC, R2.2 = R2.2, R2.1 = R2.1, 
                                                                maxT = maxT, rho.0_lev1 = rho.0_lev1,
                                                                rho.0_lev2 = rho.0_lev2, rho.1_lev2 = rho.1_lev2, 
                                                                check = check, storage_path = storage_path, 
                                                                validation_name = validation_name, runSim = runSim, 
                                                                runPump = runPump)

pander::pandoc.table(blocked_i1_2c_compare_results, paste0("PowerSimulation_", "rho_", rho,"_falseNull_",M), 
                     style = "grid", split.tables = 100)
```

### n.j = 50, M = 3, numFalse = M, J = 20,rho = (0.5, 0.5, 0.5), MDES = (0.125, 0.125, 0.125), R2 = (0,0,0)

```{r rho0.5numFalseMSpecs03, warning = FALSE, message = FALSE}
# ICC needs to be zero for constant and fixed effets
design <- c("Blocked_i1_2c")
procs <- c("Bonferroni", "BH", "Holm") 
M <- 3 
MDES <- rep(0.125, M) 
rho <- c(0.5)
ncl <-2 
S = 2000 # Number of samples for Monte Carlo Simulation.
B = 2 # Number of samples for WestFall-Young. The equivalent is snum in our new method.
tnum = 10000 # Number of test statistics (samples) for all procedures other than Westfall-Young. The default is set at 10,000.
J = 20 
n.j = 50 
p.j <- 0.5 
p.j.range <- rep(0.5,J) 
theta <- matrix(0,M,M) 
diag(theta) <- 0 
omega <- rep(0,M) 
Gamma.00 <- rep(0,M) 
sig.sq <- rep(1,M)
alpha <- 0.05 
ICC <- rep(0, M) 
R2.2 <- rep(0,M)  
R2.1 <- rep(0,M) 
maxT <- FALSE 
check <- FALSE # Usually for checking the code. Do not need to  
storage_path <- "./Validation/data"
validation_name <- "Blocked_i1_2c"
runSim <- TRUE
runPump <- TRUE
```

`r colorize("Individual Statistical power is estimated to be around 0.8. Please check the table below for estimations of other definitions of statistical power. Across all power definitions, estimation results across the __PUMP__ package, __Monte Carlo Simulation__ results and __PowerUp__ package are about the same as detailed below.", "green")`

`r colorize("We will be adding Westfall-Young step down adjustment later on.", "red")`

```{r rho0.5numFalseMRun03, warning = FALSE, message = FALSE}
blocked_i1_2c_compare_results <- validate_power_blocked_i1_2cfr(rho = rho, ncl = ncl, procs = procs, design = design, 
                                                                M = M, MDES = MDES,
                                                                p.j = p.j,p.j.range = p.j.range, S = S, B = B, 
                                                                tnum = tnum, J = J, 
                                                                n.j = n.j, theta = theta, omega = omega, 
                                                                Gamma.00 = Gamma.00, sig.sq = sig.sq,
                                                                alpha = alpha, ICC = ICC, R2.2 = R2.2, R2.1 = R2.1, 
                                                                maxT = maxT, rho.0_lev1 = rho.0_lev1,
                                                                rho.0_lev2 = rho.0_lev2, rho.1_lev2 = rho.1_lev2, 
                                                                check = check, storage_path = storage_path, 
                                                                validation_name = validation_name, runSim = runSim, 
                                                                runPump = runPump)

pander::pandoc.table(blocked_i1_2c_compare_results, paste0("PowerSimulation_", "rho_", rho,"_falseNull_",M), 
                     style = "grid", split.tables = 100)
```

# MDES validation

In this section, we validate the derived minimum detectable effect size function across different __MTP__ (Bonferroni, Benjamini-Hocheberg and Holms) and __power definition (individual and minimum_1).__  The validation is carried out by generating power values for the specific __MTP__ and __power__ definitions for a specific __MDES__. Then, we supply the generated estimate __power__ as an input parameter to see if the specified __MDES__ is returned.

## Power: 0.65 - 0.87

Below, we have input parameters of number of outcomes, minimum detectable effect size, statistical power, significance level that would estimate a high statistical power value of above __0.65__ to __1.00__ for different definitions of power. This estimated power value is used as an input to the mdes function where we validate if the _returned_ __mdes__ value is close to the __mdes__ value that was inputted to provide the estimated power.

### M = 3, numFalse = M, J = 20, n.j  = 50, rho = (0.5, 0.5, 0.5), MDES = (0.125, 0.125, 0.125), R2 = (0,0,0)

```{r mdesCommonInputs02, message = FALSE,warning = FALSE}
M <- 3
MDES <- 0.125
numFalse <- 3
J <- 20
n.j <- 50
R2.1 <- 0
R2.2 <- 0
p <- 0.5
alpha <- 0.05
numCovar.1 <- 5
numCovar.2 <- NULL
rho <- 0.5
tnum <- 10000
snum <- 2000
updateProgress <- NULL
marginError <- 0.05
ICC <- 0
omega <- NULL
```

```{r readPowerTable01, warning= FALSE, message = FALSE, echo = FALSE}

# Reading the Power Table in For the corresponding specified parameters

  # R-Squared for level 1 covariates
  R2_1 <- rep(R2.1, M)
  R2_1 <- paste(R2_1, collapse = '')
  R2_1 <- stringr::str_remove_all(R2_1, "\\.")

  # rho: correlation between outcomes
  rho_1 <- paste(rho, collapse = '')
  rho_1 <- stringr::str_remove_all(rho_1, "\\.")
  
  # MDES: minimum detectable effect size
  MDES_1 <- rep(MDES, M)
  mdes_1 <- paste(MDES_1, collapse = '')
  mdes_1 <- stringr::str_remove_all(mdes_1, "\\.")

  simulation_results <- paste0(validation_name,"_",S, "_", "simsamples", "_", M, "_", "M","_", mdes_1,"_",
                               "MDES", "_", J, "_", "J", "_", n.j, "_", "nj","_",rho_1, "_rho_", R2_1,"_R_Squared_",
                               "comparison_results",".RDS")
  power_results <- readRDS(file = here::here("Validation/data",simulation_results))

```

#### Individual Power

`r colorize("The Bonferroni, Benjamini Hocheberg and Holms' __adjusted MDES__ and the __Target MDES__ are within an acceptable margin of error for the specified targetted power. The __adjusted MDES__ is estimated at the __corresponding__ individual power in the table below. The __corresponding__ individual power is the power at which our MDES method stops looking for a mdes value close to the target mdes.", "green")`

```{r mdesBonfIndiv01, message = FALSE, echo = FALSE}
mdes_0125_bonf_indiv <- pum::mdes_blocked_i1_2c(M = M, numFalse = numFalse,J = J, n.j = n.j, 
                                                power = power_results[which(power_results$MTP == "Bonferroni"), "pum_indiv"],
                                                power.definition = "indiv", MTP = "BF", marginError = marginError,
                                                p = p, alpha = alpha, numCovar.1 = numCovar.1, numCovar.2 = numCovar.2,
                                                R2.1 = R2.1, R2.2 = R2.2, ICC = ICC, mod.type = "constant", 
                                                rho = rho, omega = omega,tnum = tnum, snum = snum, ncl = ncl)
mdes_0125_bonf_indiv <- as.data.frame(mdes_0125_bonf_indiv) %>%
                            tibble::rownames_to_column(var = "MTP")
mdes_0125_bonf_indiv$MTP <- "Bonferroni"
mdes_0125_bonf_indiv$TargettedMDES <- MDES
colnames(mdes_0125_bonf_indiv)[4] <- "Targetted MDES"
mdes_0125_bonf_indiv$TargettedIndivPower <- power_results[which(power_results$MTP == "Bonferroni"), "pum_indiv"]
colnames(mdes_0125_bonf_indiv)[5] <- "Targetted indiv power"
mdes_0125_bonf_indiv <- mdes_0125_bonf_indiv[, c(1,2,4,3,5)]
```

```{r mdesBHIndiv01, message = FALSE, echo = FALSE}
mdes_0125_bh_indiv <- pum::mdes_blocked_i1_2c(M = M, numFalse = numFalse, J = J, n.j = n.j, 
                                              power = power_results[which(power_results$MTP == "BH"), "pum_indiv"],
                                              power.definition = "indiv", MTP = "BH", marginError = marginError,
                                              p = p, alpha = alpha, numCovar.1 = numCovar.1, numCovar.2 = numCovar.2,
                                              R2.1 = R2.1, R2.2 = R2.2, ICC = ICC, mod.type = "constant", 
                                              rho = rho, omega = omega, tnum = tnum, snum = snum, ncl = ncl)
mdes_0125_bh_indiv <- as.data.frame(mdes_0125_bh_indiv) %>%
                            tibble::rownames_to_column(var = "MTP")
mdes_0125_bh_indiv$MTP <- "Benjamini_Hocheberg"
mdes_0125_bh_indiv$TargettedMDES <- MDES
colnames(mdes_0125_bh_indiv)[4] <- "Targetted MDES"
mdes_0125_bh_indiv$TargettedIndivPower <- power_results[which(power_results$MTP == "BH"), "pum_indiv"]
colnames(mdes_0125_bh_indiv)[5] <- "Targetted indiv power"
mdes_0125_bh_indiv <- mdes_0125_bh_indiv[, c(1,2,4,3,5)]
```

```{r mdesHOIndiv01, message = FALSE, echo = FALSE}
mdes_0125_holms_indiv <- pum::mdes_blocked_i1_2c(M = M, numFalse = numFalse,J = J, n.j = n.j, 
                                                  power = power_results[which(power_results$MTP == "Holm"), "pum_indiv"],
                                                  power.definition = "indiv", MTP = "HO", marginError = marginError,
                                                  p = p, alpha = alpha, numCovar.1 = numCovar.1, numCovar.2 = numCovar.2,
                                                  R2.1 = R2.1, R2.2 = R2.2, ICC = ICC, mod.type = "constant", 
                                                  rho = rho, omega = omega,tnum = tnum, snum = snum, ncl = ncl)
mdes_0125_holms_indiv <- as.data.frame(mdes_0125_holms_indiv) %>%
                            tibble::rownames_to_column(var = "MTP")
mdes_0125_holms_indiv$MTP <- "Holms"
mdes_0125_holms_indiv$TargettedMDES <- MDES
colnames(mdes_0125_holms_indiv)[4] <- "Targetted MDES"
mdes_0125_holms_indiv$TargettedIndivPower <- power_results[which(power_results$MTP == "Holm"), "pum_indiv"]
colnames(mdes_0125_holms_indiv)[5] <- "Targetted indiv power"
mdes_0125_holms_indiv <- mdes_0125_holms_indiv[, c(1,2,4,3,5)]
```

```{r mdesCombined01, message = FALSE}
mdes_0125_indiv_combined <- rbind(mdes_0125_bonf_indiv, mdes_0125_bh_indiv)
mdes_0125_indiv_combined <- rbind(mdes_0125_indiv_combined, mdes_0125_holms_indiv)
pander::pandoc.table(mdes_0125_indiv_combined, style = "grid", split.tables = 90)
```




