---
title: "Validate Power: blocked_i1_2cfr"
author: "Kristin Porter, Zarni Htet, Kristen Hunter"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: monochrome
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float: yes
  github_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    toc: yes
---

This document lays out results for estimating statistical power and minimum detectable effect size (MDES) for:
Blocked RCT, with 2 levels, and randomization done at level 1 (individual level). This document contains models assuming constant treatment effects, fixed treatment effects, and random treatment effects.

We compare results of the derived methods against Monte Carlo Simulations & the PowerUp R package on power.

```{r knit.options, echo = FALSE}
# by default, all code chunks will cache their results
knitr::opts_chunk$set(cache = FALSE, warning = FALSE, message = FALSE, echo = FALSE)
```

```{r setup}
library(here)
library(pander)    # For good table format

# default tables to round to 3 digits
panderOptions('round', 3)

# simulation and user parameters
source(here::here("Validation/Simulations", "params.R"))
params.default <- user.params.list

# validate MDES and sample size?
validate.mdes.ss = FALSE
```

```{r universal.params}
user.params.list[['K']] <- 1
user.params.list[['ICC.3']] <- rep(0, M)
user.params.list[['omega.3']] <- 0
```


# Power Validation

In this section, we validate power results for different definitions of power and different adjustment procedures.

Default parameters:

* M = `r user.params.list[['M']]`
* J = `r user.params.list[['J']]`
* rho: $\rho$ =  `r user.params.list[['rho.default']]`
* ATE ES = `r user.params.list[['ATE_ES']]`
* R2.1: $R^2_1$ = `r user.params.list[['R2.1']]`

Parameters by model type:

* Omega: $\omega_2 = 0$ for constant effects, $\omega_2$ = `r user.params.list[['omega.s']]` for fixed and random
* ICC: $\text{ICC}_2 = 0$ for constant and fixed effects, $\text{ICC}_2$ = `r user.params.list[['ICC.2']]` for random effects

```{r sim.params}
# print(sim.params.list)
```

```{r user.params}
# print(user.params.list)
```

## Varying school size

```{r nj100}
user.params.list[['n.j']] <- 100
```

$\bar{n}$ = `r user.params.list[['nbar']]`

```{r 2c.nj100}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
blocked_i1_2c_nj100_power_plot <- gen.power.results.plot(params.file.base, design)
print(blocked_i1_2c_nj100_power_plot)
```


```{r 2f.nj100}
design <- "blocked_i1_2f"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- rep(0, M)
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
blocked_i1_2f_nj100_power_plot <- gen.power.results.plot(params.file.base, design)
print(blocked_i1_2f_nj100_power_plot)
```

```{r 2r.nj100}
design <- "blocked_i1_2r"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
blocked_i1_2r_nj100_power_plot <- gen.power.results.plot(params.file.base, design)
print(blocked_i1_2r_nj100_power_plot)
```

```{r nj75}
user.params.list[['n.j']] <- 75
```

$\bar{n}$ = `r user.params.list[['nbar']]`

```{r 2c.nj75}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_nj75_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_nj75_power_plot)
```


```{r 2f.nj75}
design <- "blocked_i1_2f"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_nj75_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_nj75_power_plot)
```

```{r 2r.nj75}
design <- "blocked_i1_2r"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
blocked_i1_2r_nj75_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2r_nj75_power_plot)
```

```{r nj50}
user.params.list[['n.j']] <- 50
```

$\bar{n}$ = `r user.params.list[['nbar']]`


```{r 2c.nj50}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_nj50_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_nj50_power_plot)
```

```{r 2f.nj50}
design <- "blocked_i1_2f"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2f_nj50_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2f_nj50_power_plot)
```

```{r 2r.nj50}
design <- "blocked_i1_2r"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
blocked_i1_2r_nj50_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2r_nj50_power_plot)
```


## Varying R2

```{r r2.10.6}
user.params.list[['R2.1']] <- rep(0.6, M)  
```

$R^2_1 =$ `r user.params.list[['R2.1']]`

```{r 2c.r2.10.6}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_r2.1_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_r2.1_power_plot)
user.params.list[['R2.1']] <- params.default[['R2.1']]  
```


```{r r2.20.6}
user.params.list[['R2.2']] <- rep(0.6, M)  
```

$R^2_2 =$ `r user.params.list[['R2.2']]`
(This should not affect power!)

```{r 2c.r2.20.6}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_r2.2_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_r2.2_power_plot)
user.params.list[['R2.2']] <- params.default[['R2.2']]  
```


## Varying rho

```{r rho0.2}
rho.default <- 0.2
default.rho.matrix <- gen_corr_matrix(M = M, rho.scalar = rho.default)
user.params.list[['rho.default']] <- rho.default
user.params.list[['rho.X']] <- user.params.list[['rho.C']] <-rho.default
user.params.list[['rho.u']] <- user.params.list[['rho.v']] <- user.params.list[['rho.r']] <- rho.default
```

$\rho =$ `r user.params.list[['rho.default']]`

```{r 2c.rho0.2}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_rho0.2_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_rho0.2_power_plot)
```

```{r rho0.8}
rho.default <- 0.8
default.rho.matrix <- gen_corr_matrix(M = M, rho.scalar = rho.default)
user.params.list[['rho.default']] <- rho.default
user.params.list[['rho.X']] <- user.params.list[['rho.C']] <-rho.default
user.params.list[['rho.u']] <- user.params.list[['rho.v']] <- user.params.list[['rho.r']] <-rho.default
```

$\rho =$ `r user.params.list[['rho.default']]`

```{r 2c.rho0.8}
design <- "blocked_i1_2c"
blocked_i1_2c_rho0.8_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_rho0.8_power_plot)
# reset
rho.default <- params.default[['rho.default']]
default.rho.matrix <- gen_corr_matrix(M = M, rho.scalar = rho.default)
user.params.list[['rho.default']] <- rho.default
user.params.list[['rho.X']] <- user.params.list[['rho.C']] <-rho.default
user.params.list[['rho.u']] <- user.params.list[['rho.v']] <- user.params.list[['rho.r']] <-rho.default
```


## Varying true positives

```{r MDES0.12500}
user.params.list[['ATE_ES']] <- c(0.125, 0, 0)
```

ATE (ES) = `r user.params.list[['ATE_ES']]`

```{r 2c.MDES0.12500}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
user.params.list[['ICC.2']] <- rep(0, M)
blocked_i1_2c_mdes0.1250.00.0.00_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2c_mdes0.1250.00.0.00_power_plot)
user.params.list[['ATE_ES']] <- params.default[['ATE_ES']]
```

## Varying ICC

```{r ICC20.9}
user.params.list[['ICC.2']] <- rep(0.8, M)
```

$\text{ICC}_2$ = `r user.params.list[['ICC.2']]`

```{r 2r.ICC20.8}
design <- "blocked_i1_2r"
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
blocked_i1_2r_ICC2.0.5_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(blocked_i1_2r_ICC2.0.5_power_plot)
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
```

# MDES validation

In this section, we validate the minimum detectable effect size (MDES) across different __MTP__.  The validation is carried out by generating power values for the specific __MTP__ and __power__ definitions for a specific __MDES__. Then, we supply the generated estimate __power__ as an input parameter to see if the specified __MDES__ is returned.

```{r}
# reset to defaults
user.params.list <- params.default 
```


```{r 2c.mdesCombined}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] <- 0
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  blocked_i1_2c_mdes_results_nj50 <- readRDS(mdes.file)
  pander::pandoc.table(
    blocked_i1_2c_mdes_results_nj50,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )
}
```

```{r 2f.mdesCombined}
design <- "blocked_i1_2f"
user.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  blocked_i1_2f_mdes_results_nj50 <- readRDS(mdes.file)
  pander::pandoc.table(
    blocked_i1_2f_mdes_results_nj50,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )    
}
```

```{r 2r.mdesCombined}
design <- "blocked_i1_2r"
user.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  blocked_i1_2r_mdes_results_nj50 <- readRDS(mdes.file)
  pander::pandoc.table(
    blocked_i1_2r_mdes_results_nj50,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )
}
```

# Sample size validation

In this section, we validate the derived sample size across different __MTP__. The validation is carried out by generating power values for the specific __MTP__ and __power__ definitions for a specific __MDES__. Then, we supply the generated estimate __power__ and MDES as an input parameter to see if the specified sample size is returned.

```{r 2c.sampleCombined}
design <- "blocked_i1_2c"
user.params.list[['omega.2']] = 0
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  blocked_i1_2c_sample_results_nj50 <- readRDS(sample.file)
  pander::pandoc.table(
    blocked_i1_2c_sample_results_nj50,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```


```{r 2f.sampleCombined}
design <- "blocked_i1_2f"
user.params.list[['omega.2']] = 0.5
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  blocked_i1_2f_sample_results_nj50 <- readRDS(sample.file)
  pander::pandoc.table(
    blocked_i1_2f_sample_results_nj50,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2r.sampleCombined}
design <- "blocked_i1_2r"
user.params.list[['omega.2']] = 0.5
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  blocked_i1_2r_sample_results_nj50 <- readRDS(sample.file)
  pander::pandoc.table(
    blocked_i1_2r_sample_results_nj50,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```