---
title: "Validate Power: d3.1_m3rr2rr"
author: "Kristin Porter, Zarni Htet, Kristen Hunter"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: monochrome
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float: yes
  github_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    toc: yes
---

Design: Blocked RCT, with 3 levels, and randomization done at level 1 (individual level).

Models: random treatment effects.

```{r knit.options, echo = FALSE}
# by default, all code chunks will cache their results
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, echo = FALSE)
```

```{r setup}
library(here)
library(pander)    # For good table format

# default tables to round to 3 digits
panderOptions('round', 3)

# simulation and user parameters
source(here::here("Validation/Simulations", "params.R"))
```

```{r}
# set document-level parameters
design <- "d3.1_m3rr2rr"

# for sufficient power and stability
user.params.list[['K']] <- 15
user.params.list[['J']] <- 30
user.params.list[['nbar']] <- 100

# what we assume for validation
sim.params.list[['S']] <- 1000

params.default <- user.params.list
```


* M = `r user.params.list[['M']]`
* J = `r user.params.list[['J']]`
* K = `r user.params.list[['K']]`
* $\bar{n} = 100$ (unless otherwise noted)
* rho: $\rho$ =  `r user.params.list[['rho.default']]`
* ATE ES = `r user.params.list[['ATE_ES']]`
* R2: $R^2_1$ = `r user.params.list[['R2.1']]`, $R^2_2$ = `r user.params.list[['R2.2']]`, $R^2_3$ = `r user.params.list[['R2.3']]`
* ICC: $\text{ICC}_2$ = `r user.params.list[['ICC.2']]`, $\text{ICC}_3$ = `r user.params.list[['ICC.3']]`
* Omega: $\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

# Power Validation

## Base case

```{r 3r.nbar100}
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
d3.1_m3rr2rr_nbar100_power_plot <- gen.power.results.plot(params.file.base, design)
print(d3.1_m3rr2rr_nbar100_power_plot)
```

## Varying school size

```{r nbar75}
user.params.list[['nbar']] <- 75
```

$\bar{n}$ = `r user.params.list[['nbar']]`

```{r 3r.nbar75}
d3.1_m3rr2rr_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_nbar75_power_plot)
```

```{r nbar50}
user.params.list[['nbar']] <- 50
```

$\bar{n}$ = `r user.params.list[['nbar']]`

```{r 3r.nbar50}
d3.1_m3rr2rr_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_nbar50_power_plot)
```

```{r nreset}
user.params.list[['nbar']] <- 100
```

## Varying R2

```{r r2.10.6}
user.params.list[['R2.1']] <- rep(0.6, M)  
```

$R^2_1 =$ `r user.params.list[['R2.1']]`

```{r 3r.r2.10.6}
d3.1_m3rr2rr_r2.1_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_r2.1_power_plot)
user.params.list[['R2.1']] <- params.default[['R2.1']]  
```

```{r r2.20.6}
user.params.list[['R2.2']] <- rep(0.6, M)  
```

$R^2_2 =$ `r user.params.list[['R2.2']]`

```{r 3r.r2.20.6}
d3.1_m3rr2rr_r2.2_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_r2.2_power_plot)
user.params.list[['R2.2']] <- params.default[['R2.2']]  
```

```{r r2.30.6}
user.params.list[['R2.3']] <- rep(0.6, M)  
```

$R^2_3 =$ `r user.params.list[['R2.3']]`

```{r 3r.r2.30.6}
d3.1_m3rr2rr_r2.3_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_r2.3_power_plot)
user.params.list[['R2.3']] <- params.default[['R2.3']]  
```

```{r r2.10r2.20r2.30}
user.params.list[['R2.1']] <- rep(0, M)  
user.params.list[['R2.2']] <- rep(0, M)  
user.params.list[['R2.3']] <- rep(0, M)  
```

$R^2_1 =$ `r user.params.list[['R2.1']]`
$R^2_2 =$ `r user.params.list[['R2.2']]`
$R^2_3 =$ `r user.params.list[['R2.3']]`

```{r 3r.r2.10r2.20r2.30}
d3.1_m3rr2rr_r2.3_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_r2.3_power_plot)
user.params.list[['R2.1']] <- params.default[['R2.1']]  
user.params.list[['R2.2']] <- params.default[['R2.2']]  
user.params.list[['R2.3']] <- params.default[['R2.3']]  
```


## Varying rho

```{r rho0.2tPMnbar50}
rho.default <- 0.2
default.rho.matrix <- gen_corr_matrix(M = M, rho.scalar = rho.default)
user.params.list[['rho.default']] <- rho.default
user.params.list[['rho.X']] <- user.params.list[['rho.C']] <- rho.default
user.params.list[['rho.u']] <- user.params.list[['rho.v']] <- user.params.list[['rho.r']] <- rho.default
```

$\rho =$ `r user.params.list[['rho.default']]`

```{r 3r.rho0.2}
d3.1_m3rr2rr_rho0.2_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_rho0.2_power_plot)
```

```{r rho0.8}
rho.default <- 0.8
default.rho.matrix <- gen_corr_matrix(M = M, rho.scalar = rho.default)
user.params.list[['rho.default']] <- rho.default
user.params.list[['rho.X']] <- user.params.list[['rho.C']] <-rho.default
user.params.list[['rho.u']] <- user.params.list[['rho.v']] <- user.params.list[['rho.r']] <-rho.default
```

$\rho =$ `r user.params.list[['rho.default']]`

```{r 3r.rho0.8}
d3.1_m3rr2rr_rho0.8_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_rho0.8_power_plot)
# reset
rho.default <- params.default[['rho.default']]
default.rho.matrix <- gen_corr_matrix(M = M, rho.scalar = rho.default)
user.params.list[['rho.default']] <- rho.default
user.params.list[['rho.X']] <- user.params.list[['rho.C']] <-rho.default
user.params.list[['rho.u']] <- user.params.list[['rho.v']] <- user.params.list[['rho.r']] <-rho.default
```


## Varying true positives

```{r MDES0.12500}
user.params.list[['ATE_ES']] <- c(0.125, 0, 0)
```

ATE (ES) = `r user.params.list[['ATE_ES']]`

```{r 23r.MDES0.12500}
d3.1_m3rr2rr_mdes0.1250.00.0.00_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_mdes0.1250.00.0.00_power_plot)
user.params.list[['ATE_ES']] <- params.default[['ATE_ES']]
```

## Varying ICC

```{r ICC20.7}
user.params.list[['ICC.2']] <- rep(0.7, M)
```

$\text{ICC}_2$ = `r user.params.list[['ICC.2']]`

```{r 3r.ICC20.7}
d3.1_m3rr2rr_ICC2.0.7_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_ICC2.0.7_power_plot)
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
```

```{r ICC30.7}
user.params.list[['ICC.3']] <- rep(0.7, M)
```

$\text{ICC}_3$ = `r user.params.list[['ICC.3']]`

```{r 3r.ICC30.7}
d3.1_m3rr2rr_ICC3.0.7_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_ICC3.0.7_power_plot)
user.params.list[['ICC.3']] <- params.default[['ICC.3']]
```


```{r ICC20}
user.params.list[['ICC.2']] <- rep(0, M)
```

$\text{ICC}_2$ = `r user.params.list[['ICC.2']]`

```{r 3r.ICC20}
d3.1_m3rr2rr_ICC2.0_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_ICC2.0_power_plot)
user.params.list[['ICC.2']] <- params.default[['ICC.2']]
```

```{r ICC30}
user.params.list[['ICC.3']] <- rep(0, M)
```

$\text{ICC}_2$ = `r user.params.list[['ICC.2']]`

```{r 3r.ICC30}
d3.1_m3rr2rr_ICC3.0_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_ICC3.0_power_plot)
user.params.list[['ICC.3']] <- params.default[['ICC.3']]
```

## Varying Omega

```{r omega20.8}
user.params.list[['omega.2']] <- 0.8
user.params.list[['omega.3']] <- params.default[['omega.3']]
```

$\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

```{r 3r.omega20.8}
d3.1_m3rr2rr_omega2.0.8_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_omega2.0.8_power_plot)
```

```{r omega30.8}
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['omega.3']] <- 0.8
```

$\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

```{r 3r.omega30.8}
d3.1_m3rr2rr_omega3.0.8_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_omega3.0.8_power_plot)
```

```{r omega20.8omega30.8}
user.params.list[['omega.2']] <- 0.8
user.params.list[['omega.3']] <- 0.8
```

$\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

```{r 3r.omega20.8omega30.8}
d3.1_m3rr2rr_omega2.0.8omega3.0.8_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_omega2.0.8omega3.0.8_power_plot)
user.params.list[['omega.3']] <- params.default[['omega.3']]
user.params.list[['omega.2']] <- params.default[['omega.2']]
```


```{r omega20}
user.params.list[['omega.2']] <- 0
user.params.list[['omega.3']] <- params.default[['omega.3']]
```

$\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

```{r 3r.omega20}
d3.1_m3rr2rr_omega2.0_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_omega2.0_power_plot)
user.params.list[['omega.2']] <- params.default[['omega.2']]
```


```{r omega30}
user.params.list[['omega.3']] <- 0
```

$\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

```{r 3r.omega30}
d3.1_m3rr2rr_omega3.0_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_omega3.0_power_plot)
user.params.list[['omega.3']] <- params.default[['omega.3']]
```

```{r omega2030}
user.params.list[['omega.2']] <- 0
user.params.list[['omega.3']] <- 0
```

$\omega_2$ = `r user.params.list[['omega.2']]`, $\omega_3$ = `r user.params.list[['omega.3']]`

```{r 3r.omega2030}
d3.1_m3rr2rr_omega2.03.0_power_plot <- gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_omega2.03.0_power_plot)
user.params.list[['omega.2']] <- params.default[['omega.2']]
user.params.list[['omega.3']] <- params.default[['omega.3']]
```

## Kappa

```{r kappa0.4}
kappa <- 0.4
kappa.matrix <- diag(kappa, M)
user.params.list[['kappa.w']] <- user.params.list[['kappa.u']] <- kappa.matrix
```

$\kappa$ = `r kappa`

```{r 3r.kappa0.4}
design <- "d3.1_m3rr2rr"
d3.1_m3rr2rr_kappa0.4_power_plot <-
  gen.power.results.plot(gen_params_file_base(user.params.list, sim.params.list, design), design)
print(d3.1_m3rr2rr_kappa0.4_power_plot)
```

```{r kappa0.4reset}
user.params.list[['kappa.w']] <- user.params.list[['kappa.u']] <- params.default[['kappa.w']]
```

# MDES validation

```{r}
# reset to defaults
user.params.list <- params.default 
```

```{r 3r.mdes}
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d3.1_m3rr2rr_mdes_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d3.1_m3rr2rr_mdes_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )
}
```

# Sample size validation

```{r 2r.sampleD1.J}
design <- "d3.1_m3rr2rr"
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d3.1_m3rr2rr_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d3.1_m3rr2rr_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2r.sampleD1.K}
design <- "d3.1_m3rr2rr"
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_K_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d3.1_m3rr2rr_sample_D1_K_results <- readRDS(sample.file)
  pander::pandoc.table(
    d3.1_m3rr2rr_sample_D1_K_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2r.sampleD1.nbar}
design <- "d3.1_m3rr2rr"
params.file.base <- gen_params_file_base(user.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_nbar_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d3.1_m3rr2rr_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d3.1_m3rr2rr_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```
