---
title: "Validation Template for Design X"
output: github_document

---

# Sourcing all needed code 

```{r chunk1_source }
# source functions for computing power, MDES, sample size that correspond to design of focus
# source functions for simulations that correspond to design of focus
```


# Design and parameter specifications

```{r chunk2_specifications}
ncl<-24
rho<-c(0.2,0.5,0.8)
M <- c(1,3,6)
sigma<-matrix(rep(rho,M*M),nrow=M,ncol=M)
diag(sigma)<-1
mdes_<-c(0.125,0.5)
MDES1<-c(rep(mdes_,M))  #this doesnt run, M has too many arguments- loop?
MDES2<-c(rep(mdes_,M*2/3),rep(0,M/3))
MDES3<-c(rep(mdes_,M/3),rep(0,M*2/3))
p <- 0.5
R2.1 <- c(0,0.5)
R2.2 <- c(0,0.5)
numCovar.1 <- c(0,1)
numCovar.2 <- c(0,1)
ICC <- c(0,0.5)
J <- c(20,50)
n.j <- c(20,50)
alpha = 0.05
```

# Verify power calculations and capture run times

## Maybe different chunk for each design specification (e.g, constant, fixed, random effects)

### The simulations will take a long time so may want to run these separately, and pull results into this markdown??

```{r chunk3_verifyPower}
# loop through combinations (don't need all crossed - TBD)

  # run new

  # run simulations

  # for M=1, look up in Power-Up

  # create table of results and check for if within TBD margin of error

  # create table of run times

```

# Verify MDES calculations and capture run times

```{r chunk4_verifyMDES}

# Specify power estimates that match those obtained in chunk3_verifyPower (grab from final table/matrix) and corresponding sample size - and check that returned MDES is correct

```

# Verify sample size calculations and capture run times

```{r chunk4_verifySampleSize}

# Specify power estimates that match those obtained in chunk3_verifyPower (grab from final table/matrix) and corresponding MDES - and check that returned sample size is correct

# repeat for number of blocks, number within blocks, etc. 

```

# ?? Verify error messages

```{r chunk4_verifyErrorsWarnings}

# Pass in anticipated mistakes and show return of error messages/warnings?

```

# Zarni's edits

### Libraries and Source Files

```{r}
#
```

### Parameters for Monte Carlo Simulation Runs

#### Subtitle: Type of Function we are validating

```{r}

```

### Firing Up & Storing Results from Monte Carlo Simulation Runs

```{r}

```

### Firing up & Storing Results from derived function

```{r}

```

### Comparing results for validation

```{r}

```




