---
title: "Validate Power: d2.1_m2fcfr"
author: "Kristin Porter, Zarni Htet, Kristen Hunter"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    highlight: monochrome
    theme: spacelab
    toc: yes
    toc_depth: 5
    toc_float: yes
  github_document:
    df_print: paged
    fig_caption: yes
    fig_height: 8
    fig_width: 10
    toc: yes
---

Design: Blocked RCT, with 2 levels, and randomization done at level 1 (individual level).

Models: Constant treatment effects, fixed treatment effects, and random treatment effects.

```{r knit.options, echo = FALSE}
knitr::opts_chunk$set(cache = FALSE, warning = FALSE, message = FALSE, echo = FALSE)
```

```{r setup}
library(PUMP)
library(here)
library(pander)    # For good table format
# default tables to round to 3 digits
panderOptions('round', 3)
# simulation and user parameters
source(here::here("Validation/Simulations", "params.R"))
```

```{r universal.params}
# assumptions
model.params.list[['K']] <- 1
model.params.list[['J']] <- 20
model.params.list[['ICC.3']] <- 0
model.params.list[['omega.3']] <- 0
model.params.list[['R2.3']] <- 0
# what we assume for validation
sim.params.list[['S']] <- 10000
params.default <- model.params.list
```

Default parameters:

* M = `r model.params.list[['M']]`
* J = `r model.params.list[['J']]`
* rho: $\rho$ =  `r model.params.list[['rho.default']]`
* ATE ES = `r model.params.list[['ATE_ES']]`
* R2.1: $R^2_1$ = `r model.params.list[['R2.1']]`

Parameters by model type:

* Omega: $\omega_2 = 0$ for constant effects, $\omega_2$ = `r model.params.list[['omega.2']]` for fixed and random
* ICC: $\text{ICC}_2 = 0$ for constant and fixed effects, $\text{ICC}_2$ = `r model.params.list[['ICC.2']]` for random effects

Assumptions:

* Two-level design: $\text{ICC}_3 = 0$,  $\omega_3 = 0$, $K = 1$

# Power Validation

## Base case

```{r 2c.nbar50}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_nbar50_power_plot)
```

```{r 2f.nbar50}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2ff_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2ff_nbar50_power_plot)
```

```{r 2r.nbar50}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fr_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_nbar50_power_plot)
```

## Varying school size

```{r nbar100}
model.params.list[['nbar']] <- 100
```

$\bar{n}$ = `r model.params.list[['nbar']]`

```{r 2c.nbar100}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
# model.params.list[['ICC.2']] <- rep(0, M)
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
d2.1_m2fc_nbar100_power_plot <- gen.power.results.plot(params.file.base, design)
print(d2.1_m2fc_nbar100_power_plot)
```


```{r 2f.nbar100}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
# model.params.list[['ICC.2']] <- rep(0, M)
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
d2.1_m2ff_nbar100_power_plot <- gen.power.results.plot(params.file.base, design)
print(d2.1_m2ff_nbar100_power_plot)
```

```{r 2r.nbar100}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
d2.1_m2fr_nbar100_power_plot <- gen.power.results.plot(params.file.base, design)
print(d2.1_m2fr_nbar100_power_plot)
```

```{r nbar75}
model.params.list[['nbar']] <- 75
```

$\bar{n}$ = `r model.params.list[['nbar']]`

```{r 2c.nbar75}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_nbar75_power_plot)
```


```{r 2f.nbar75}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_nbar75_power_plot)
```

```{r 2r.nbar75}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fr_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_nbar75_power_plot)
```

```{r nbar50}
model.params.list[['nbar']] <- 50
```

## Varying R2

```{r r2.10.6}
model.params.list[['R2.1']] <- rep(0.6, M)  
```

$R^2_1 =$ `r model.params.list[['R2.1']]`

```{r 2c.r2.10.6}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_r2.1_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_r2.1_power_plot)
model.params.list[['R2.1']] <- params.default[['R2.1']]  
```


```{r r2.20.6}
model.params.list[['R2.2']] <- rep(0.6, M)  
```

$R^2_2 =$ `r model.params.list[['R2.2']]`

```{r 2c.r2.20.6}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_r2.2_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_r2.2_power_plot)
model.params.list[['R2.2']] <- params.default[['R2.2']]  
```


```{r r2.10r2.20}
model.params.list[['R2.1']] <- rep(0, M)  
model.params.list[['R2.2']] <- rep(0, M)  
```

$R^2_1 =$ `r model.params.list[['R2.2']]`
$R^2_2 =$ `r model.params.list[['R2.1']]`

```{r 2c.r2.10r2.20}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_r2.2_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_r2.2_power_plot)
model.params.list[['R2.1']] <- params.default[['R2.1']]  
model.params.list[['R2.2']] <- params.default[['R2.2']]  
```

## Varying rho

```{r rho0.2}
rho.default <- 0.2
default.rho.matrix <- PUMP::gen_corr_matrix(M = M, rho.scalar = rho.default)
model.params.list[['rho.default']] <- rho.default
model.params.list[['rho.X']] <- model.params.list[['rho.C']] <-rho.default
model.params.list[['rho.u']] <- model.params.list[['rho.v']] <- model.params.list[['rho.r']] <- rho.default
```

$\rho =$ `r model.params.list[['rho.default']]`

```{r 2c.rho0.2}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_rho0.2_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_rho0.2_power_plot)
```

```{r rho0.8}
rho.default <- 0.8
default.rho.matrix <- PUMP::gen_corr_matrix(M = M, rho.scalar = rho.default)
model.params.list[['rho.default']] <- rho.default
model.params.list[['rho.X']] <- model.params.list[['rho.C']] <-rho.default
model.params.list[['rho.u']] <- model.params.list[['rho.v']] <- model.params.list[['rho.r']] <-rho.default
```

$\rho =$ `r model.params.list[['rho.default']]`

```{r 2c.rho0.8}
design <- "d2.1_m2fc"
d2.1_m2fc_rho0.8_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_rho0.8_power_plot)
# reset
rho.default <- params.default[['rho.default']]
default.rho.matrix <- PUMP::gen_corr_matrix(M = M, rho.scalar = rho.default)
model.params.list[['rho.default']] <- rho.default
model.params.list[['rho.X']] <- model.params.list[['rho.C']] <-rho.default
model.params.list[['rho.u']] <- model.params.list[['rho.v']] <- model.params.list[['rho.r']] <-rho.default
```


## Varying true positives

```{r MDES0.12500}
model.params.list[['ATE_ES']] <- c(0.125, 0, 0)
```

ATE (ES) = `r model.params.list[['ATE_ES']]`

```{r 2c.MDES0.12500}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fc_mdes0.1250.00.0.00_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_mdes0.1250.00.0.00_power_plot)
model.params.list[['ATE_ES']] <- params.default[['ATE_ES']]
```

## Varying ICC

Note: we expect a discrepancy when ICC is not zero between powerup and pump.

```{r ICC20.7}
model.params.list[['ICC.2']] <- rep(0.7, M)
```

$\text{ICC}_2$ = `r model.params.list[['ICC.2']]`

```{r 2c.ICC20.7}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
d2.1_m2fc_ICC2.0.7_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_ICC2.0.7_power_plot)
```

```{r 2f.ICC20.7}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2ff_ICC2.0.7_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2ff_ICC2.0.7_power_plot)
```

```{r 2r.ICC20.7}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
d2.1_m2fr_ICC2.0.7_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_ICC2.0.7_power_plot)
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
```


```{r ICC20}
model.params.list[['ICC.2']] <- rep(0, M)
```

$\text{ICC}_2$ = `r model.params.list[['ICC.2']]`

```{r 2c.ICC20}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
d2.1_m2fc_ICC2.0_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_ICC2.0_power_plot)
```

```{r 2f.ICC20}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2ff_ICC2.0_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2ff_ICC2.0_power_plot)
```

```{r 2r.ICC20}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2fr_ICC2.0_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_ICC2.0_power_plot)
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
```

## Varying Omega

```{r omega0.8}
model.params.list[['omega.2']] <- 0.8
```

$\omega_2$ = `r model.params.list[['omega.2']]`

```{r 2f.omega20.8}
design <- "d2.1_m2ff"
d2.1_m2ff_omega20.8_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2ff_omega20.8_power_plot)
```

```{r 2r.omega0.8}
design <- "d2.1_m2fr"
d2.1_m2fr_omega20.8_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_omega20.8_power_plot)
```


```{r omega20}
model.params.list[['omega.2']] <- 0
```

$\omega_2$ = `r model.params.list[['omega.2']]`

```{r 2f.omega20}
design <- "d2.1_m2ff"
d2.1_m2ff_omega20_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2ff_omega20_power_plot)
```

```{r 2r.omega0}
design <- "d2.1_m2fr"
d2.1_m2fr_omega20_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_omega20_power_plot)
```

```{r omega0reset}
model.params.list[['omega.2']] <- params.default[['omega.2']]
```

## Kappa

```{r kappa0.4}
kappa <- 0.4
kappa.matrix <- diag(kappa, M)
model.params.list[['kappa.w']] <- model.params.list[['kappa.u']] <- kappa.matrix
```

$\kappa$ = `r kappa`

```{r 2c.kappa0.4}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
d2.1_m2fc_kappa0.4_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fc_kappa0.4_power_plot)
```

```{r 2f.kappa0.4}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2ff_kappa0.4_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2ff_kappa0.4_power_plot)
```

```{r 2r.kappa0.4}
design <- "d2.1_m2fr"
d2.1_m2fr_kappa0.4_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, design), design)
print(d2.1_m2fr_kappa0.4_power_plot)
```

```{r kappa0.4reset}
model.params.list[['kappa.w']] <- model.params.list[['kappa.u']] <- params.default[['kappa.w']]
```

# MDES validation

```{r}
model.params.list <- params.default 
```


```{r 2c.mdes.D1}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fc_mdes_D1_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d2.1_m2fc_mdes_D1_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )
}
```

```{r 2f.mdes.D1}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2ff_mdes_D1_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d2.1_m2ff_mdes_D1_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )    
}
```

```{r 2r.mdes.D1}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fr_mdes_D1_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d2.1_m2fr_mdes_D1_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = design
  )
}
```

# Sample size validation

```{r 2c.sampleD1.J}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fc_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fc_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2c.sampleD1.nbar}
design <- "d2.1_m2fc"
model.params.list[['omega.2']] <- 0
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_nbar_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fc_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fc_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2f.sampleD1.J}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2ff_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2ff_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2f.sampleD1.nbar}
design <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_nbar_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2ff_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2ff_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```


```{r 2r.sampleD1.J}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fr_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fr_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```

```{r 2r.sampleD1.nbar}
design <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, design)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fr_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fr_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = design
  )
}
```