---
title: "Validate Power: d2.1"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

d_m: Blocked RCT, with 2 levels, and randomization done at level 1 (individual level).

Models: Constant treatment effects, fixed treatment effects, and random treatment effects.

Note: we expect a discrepancy when ICC is not zero between powerup and pump.

```{r knit.options, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, warning = FALSE, message = FALSE, echo = FALSE,
  fig.height = 7
)
```

```{r setup}
library(here)
library(PUMP)
library(pander)
# default tables to round to 3 digits
panderOptions('round', 3)
source(here::here("validation/code", "misc.R"))
```

```{r sim.params}
sim.params.list <- list(
  S = 5000                     # Number of samples for Monte Carlo Simulation
  , Q = 1
)
```

```{r model.params}
M <- 3
model.params.list <- list(
  M = 3                                   # number of outcomes
  , J = 20                                # number of schools
  , K = 1                                 # number of districts (for two-level model, set K = 1)
  , nbar = 50                             # number of individuals per school
  , rho.default = 0.5                     # default rho value
  ################################################## grand mean outcome and impact
  , MDES = rep(0.125, M)                  # minimum detectable effect size      
  ################################################## level 2: schools
  , ICC.2 = rep(0.2, M)                   # school intraclass correlation	
  , omega.2 = rep(0.1, M)                 # ratio of school effect size variability to random effects variability
  ################################################## level 1: individuals
  , numCovar.1 = 1                        # number of individual covariates
  , R2.1 = rep(0.1, M)                    # percent of indiv variation explained by indiv covariates
)

params.default <- model.params.list
```

Default parameters:

* M = `r model.params.list[['M']]`
* J = `r model.params.list[['J']]`
* $\rho$ =  `r model.params.list[['rho.default']]`
* MDES = `r model.params.list[['MDES']]`
* $R^2_1$ = `r model.params.list[['R2.1']]`
* $\text{ICC}_2$ = `r model.params.list[['ICC.2']]`

Parameters by model type:

* Omega: $\omega_2 = 0$ for constant effects, $\omega_2$ = `r model.params.list[['omega.2']]` for fixed and random

Assumptions:

* Two-level design: $\text{ICC}_3 = 0$,  $\omega_3 = 0$, $K = 1$

*Remark*. For all $d2.1$ designs and models, PowerUpR! assumes $ICC.2 = 0$, but we do not make that assumption here. Thus, we expect to see a discrepancy betweeen PUMP and Powerup except for the setting when we assume $ICC.2 = 0$.

# Power Validation

## Base case

```{r 2c_nbar50}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_nbar50_power_plot)
```

```{r 2f_nbar50}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2ff_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2ff_nbar50_power_plot)
```

```{r 2r_nbar50}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2fr_nbar50_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fr_nbar50_power_plot)
```

\newpage
## Varying school size

```{r nbar100}
model.params.list[['nbar']] <- 100
```

$\bar{n}$ = `r model.params.list[['nbar']]`

```{r 2c_nbar100}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
d2.1_m2fc_nbar100_power_plot <- gen.power.results.plot(params.file.base, d_m)
print(d2.1_m2fc_nbar100_power_plot)
```


```{r 2f_nbar100}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
d2.1_m2ff_nbar100_power_plot <- gen.power.results.plot(params.file.base, d_m)
print(d2.1_m2ff_nbar100_power_plot)
```

```{r 2r_nbar100}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
d2.1_m2fr_nbar100_power_plot <- gen.power.results.plot(params.file.base, d_m)
print(d2.1_m2fr_nbar100_power_plot)
```
\newpage
```{r nbar75}
model.params.list[['nbar']] <- 75
```

$\bar{n}$ = `r model.params.list[['nbar']]`

```{r 2c_nbar75}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_nbar75_power_plot)
```


```{r 2f_nbar75}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2fc_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_nbar75_power_plot)
```

```{r 2r_nbar75}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2fr_nbar75_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fr_nbar75_power_plot)
```

```{r nbar50}
model.params.list[['nbar']] <- 50
```

\newpage
## Varying R2

```{r r2_106}
model.params.list[['R2.1']] <- rep(0.6, M)  
```

$R^2_1 =$ `r model.params.list[['R2.1']]`

```{r 2c_r2_106}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_r2.1_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_r2.1_power_plot)
model.params.list[['R2.1']] <- params.default[['R2.1']]  
```

\newpage
```{r r2_10r2_20}
model.params.list[['R2.1']] <- rep(0, M)  
```

$R^2_1 =$ `r model.params.list[['R2.1']]`

```{r 2c_r2_10r2_20}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_r2.2_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_r2.2_power_plot)
model.params.list[['R2.1']] <- params.default[['R2.1']]  
```

\newpage
## Varying rho

```{r rho0_2}
rho.default <- 0.2
default.rho.matrix <- PUMP::gen_corr_matrix(M = M, rho.scalar = rho.default)
model.params.list[['rho.default']] <- rho.default
model.params.list[['rho.X']] <- model.params.list[['rho.C']] <-rho.default
model.params.list[['rho.u']] <- model.params.list[['rho.v']] <- model.params.list[['rho.r']] <- rho.default
```

$\rho =$ `r model.params.list[['rho.default']]`

```{r 2c_rho0_2}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_rho0.2_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_rho0.2_power_plot)
```

```{r rho0_8}
rho.default <- 0.8
default.rho.matrix <- PUMP::gen_corr_matrix(M = M, rho.scalar = rho.default)
model.params.list[['rho.default']] <- rho.default
model.params.list[['rho.X']] <- model.params.list[['rho.C']] <-rho.default
model.params.list[['rho.u']] <- model.params.list[['rho.v']] <- model.params.list[['rho.r']] <-rho.default
```
\newpage
$\rho =$ `r model.params.list[['rho.default']]`

```{r 2c_rho0_8}
d_m <- "d2.1_m2fc"
d2.1_m2fc_rho0.8_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_rho0.8_power_plot)
# reset
rho.default <- params.default[['rho.default']]
default.rho.matrix <- PUMP::gen_corr_matrix(M = M, rho.scalar = rho.default)
model.params.list[['rho.default']] <- rho.default
model.params.list[['rho.X']] <- model.params.list[['rho.C']] <-rho.default
model.params.list[['rho.u']] <- model.params.list[['rho.v']] <- model.params.list[['rho.r']] <-rho.default
```


\newpage
## Varying true positives

```{r MDES012500}
model.params.list[['MDES']] <- c(0.125, 0, 0)
```

MDES = `r model.params.list[['MDES']]`

```{r 2c_MDES012500}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_mdes0.1250.00.0.00_power_plot <- gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_mdes0.1250.00.0.00_power_plot)
model.params.list[['MDES']] <- params.default[['MDES']]
```

\newpage
## Varying ICC



```{r ICC207}
model.params.list[['ICC.2']] <- rep(0.7, M)
```

$\text{ICC}_2$ = `r model.params.list[['ICC.2']]`

```{r 2c_ICC207}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_ICC2.0.7_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_ICC2.0.7_power_plot)
```

```{r 2f_ICC207}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2ff_ICC2.0.7_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2ff_ICC2.0.7_power_plot)
```

```{r 2r_ICC207}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2fr_ICC2.0.7_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fr_ICC2.0.7_power_plot)
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
```


```{r ICC20}
model.params.list[['ICC.2']] <- rep(0, M)
```
\newpage
$\text{ICC}_2$ = `r model.params.list[['ICC.2']]`

```{r 2c_ICC20}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
d2.1_m2fc_ICC2.0_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fc_ICC2.0_power_plot)
```

```{r}
png(file = here::here('validation/results', 'example_validation_plot.png'))
print(d2.1_m2fc_ICC2.0_power_plot)
dev.off()
```

```{r 2f_ICC20}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2ff_ICC2.0_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2ff_ICC2.0_power_plot)
```

```{r 2r_ICC20}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
d2.1_m2fr_ICC2.0_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fr_ICC2.0_power_plot)
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
```

\newpage
## Varying Omega

```{r omega08}
model.params.list[['omega.2']] <- rep(0.8, M)
```

$\omega_2$ = `r model.params.list[['omega.2']]`

```{r 2f_omega208}
d_m <- "d2.1_m2ff"
d2.1_m2ff_omega20.8_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2ff_omega20.8_power_plot)
```

```{r 2r_omega08}
d_m <- "d2.1_m2fr"
d2.1_m2fr_omega20.8_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fr_omega20.8_power_plot)
```


```{r omega20}
model.params.list[['omega.2']] <- rep(0, M)
```
\newpage
$\omega_2$ = `r model.params.list[['omega.2']]`

```{r 2f_omega20}
d_m <- "d2.1_m2ff"
d2.1_m2ff_omega20_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2ff_omega20_power_plot)
```

```{r 2r_omega0}
d_m <- "d2.1_m2fr"
d2.1_m2fr_omega20_power_plot <-
  gen.power.results.plot(gen_params_file_base(model.params.list, sim.params.list, d_m), d_m)
print(d2.1_m2fr_omega20_power_plot)
```

```{r omega0reset}
model.params.list[['omega.2']] <- params.default[['omega.2']]
```

\newpage
# MDES validation

Target value: `r model.params.list[['MDES']][1]`

```{r}
model.params.list <- params.default 
```


```{r 2c_mdes_D1}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fc_mdes_D1_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d2.1_m2fc_mdes_D1_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = d_m
  )
}
```


```{r, include = FALSE}
library(knitr)
kable(d2.1_m2ff_mdes_D1_results, format = 'latex', digits = 3)
```

```{r 2f_mdes_D1}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2ff_mdes_D1_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d2.1_m2ff_mdes_D1_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = d_m
  )    
}
```

```{r 2r_mdes_D1}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
model.params.list[['ICC.2']] <- params.default[['ICC.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
mdes.file <- find_file(params.file.base, type = 'mdes_D1indiv')
if(length(mdes.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fr_mdes_D1_results <- readRDS(mdes.file)
  pander::pandoc.table(
    d2.1_m2fr_mdes_D1_results,
    style = "grid", split.tables = 100, row.names = FALSE, 
    caption = d_m
  )
}
```

\newpage
# Sample size validation


Target value: `r model.params.list[['J']]`

```{r 2c_sampleD1_J}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fc_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fc_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = d_m
  )
}
```

```{r, include = FALSE}
kable(d2.1_m2fc_sample_D1_J_results, format = 'latex', digits = 3)
```

```{r 2c_sampleD1_nbar}
d_m <- "d2.1_m2fc"
model.params.list[['omega.2']] <- NULL
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
sample.file <- find_file(params.file.base, type = 'sample_nbar_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fc_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fc_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = d_m
  )
}
```

```{r 2f_sampleD1_J}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2ff_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2ff_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = d_m
  )
}
```


Target value: `r model.params.list[['nbar']]`

```{r 2f_sampleD1_nbar}
d_m <- "d2.1_m2ff"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
sample.file <- find_file(params.file.base, type = 'sample_nbar_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2ff_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2ff_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = d_m
  )
}
```


```{r 2r_sampleD1_J}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fr_sample_D1_J_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fr_sample_D1_J_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = d_m
  )
}
```

```{r 2r_sampleD1_nbar}
d_m <- "d2.1_m2fr"
model.params.list[['omega.2']] <- params.default[['omega.2']]
params.file.base <- gen_params_file_base(model.params.list, sim.params.list, d_m)
sample.file <- find_file(params.file.base, type = 'sample_J_D1indiv')
if(length(sample.file) == 0)
{
  warning(paste('Results not yet computed for given parameters:', params.file.base))
} else
{
  d2.1_m2fr_sample_D1_nbar_results <- readRDS(sample.file)
  pander::pandoc.table(
    d2.1_m2fr_sample_D1_nbar_results,
    style = "grid", split.tables = 100, row.names = FALSE,
    caption = d_m
  )
}
```